### Choosing the right type of visualization

## Distributions
  - Histograms
  - Density Charts

## Ranking
  - Bar plots
  - Lollipop
  - Dumbbell

## Evolutions
  - Line graphs
  - Area graph

## Numeric Relationships
 - Scatter plot
 - 2d density


1) THink about the message you want to convey

2) Consider the types of variables you're working with - and data types

3) Try multiple forms

4) Arrange the components of the graphic

5) Test the outcomes of your graphic on others



```{r}
library(tidyverse)
library(here)

```

```{r}

url <- ("https://raw.githubusercontent.com/owid/energy-data/refs/heads/master/owid-energy-data.csv")
raw_data <- read_csv(url)


useful <- c("country", "coal_share_elec", "biofuel_share_elec", "gas_share_elec", "nuclear_share_elec", "oil_share_elec", "other_renewables_share_elec", "hydro_share_elec", "solar_share_elec", "wind_share_elec", "year" )


filtered_data <- raw_data %>% 
  select(useful) %>% 
  filter(country %in% c("India", "China", "United States", "Indonesia", "Pakistan", "Nigeria", "Brazil", "Bangladesh", "Russia", "Mexico"), year == 2024)

```


```{r}




plot_data <- filtered_data %>%
  select(-year) %>% 
pivot_longer(cols = -country, 
             names_to = "energy_type", 
             values_to = "value")

ggplot(plot_data, aes(x = country, y = value, fill = energy_type)) +
  geom_col(position = "fill") +
  coord_flip() +
  theme_minimal() +
  labs(y = "Proportion") +
  scale_fill_brewer(palette = "Set3")
  

```


